Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-02-07 transmission.tex
chapter 6 comment

In this chapter, I use the preferred estimate of the duration of RT-PCR positivity (see \cref{E-imperf-test:sec:conclusion}) to estimate the transmission of SARS-CoV-2 in England.
This estimation is based only on the CIS data (see \cref{transmission:sec:data} for further details).
I will show that this data is sufficient to estimate the parameters of the transmission process, suggesting surveys of this type are sufficient for surveillance of an epidemic.

@@

\todo[inline]{This is now mostly in chapter 2, double-check there's nothing missed

@@

The deterministic relationship between incidence proportion and prevalence means that either can be inferred, with the other then calculated.
Denote the quantity to be inferred at time $t$ is strata $i$ as $\pi_{it}$.
Note that for any $i$ and $t$, the number of positives will be small, often fewer than 5; therefore some form of borrowing between time and strata is required.
The statistical model used for inference, including the borrowing of information, should have the following properties.
\begin{itemize}
    \item The smoothing model should penalise towards exponential growth.
        Penalising towards exponential growth is a natural formulation because mechanistic models predict exponential growth over short periods of time.
        Furthermore, such penalisation has previously been shown to improve estimates of prevalence by balancing over and under-fitting~\autocite{ealesAppropriately}.
\end{itemize}

>>>>

In this chapter, I use the preferred estimate of the duration of RT-PCR positivity (see \cref{E-imperf-test:sec:conclusion}) to estimate the transmission of SARS-CoV-2 in England.
This estimation is based only on the CIS data (as shown in \cref{E-biology-data:sec:positivity} for further details).
I will show that this data is sufficient to estimate the parameters of the transmission process, suggesting surveys of this type are sufficient for surveillance of an epidemic.

@@

\todo[inline]{This is now in \cref{E-biology-data:sec:posivity})}

@@

The deterministic relationship between incidence proportion and prevalence means that either can be inferred, with the other then calculated.
Denote the quantity to be inferred at time $t$ is strata $i$ as $\pi_{it}$.
Note that for any $i$ and $t$, the number of positives will be small, often fewer than 5; therefore some form of borrowing between time and strata is required.
The statistical model and inference procedure, including the borrowing of information, should have the following properties.
\begin{itemize}
    \item The smoothing model should penalise towards exponential growth.
        Penalising towards exponential growth is a natural formulation because mechanistic models predict exponential growth over short periods of time.
        Furthermore, such penalisation has previously been shown to improve estimates of prevalence by balancing over and under-fitting~\autocite{ealesAppropriately}.
    \item Limit estimates to epidemiologically and biologically plausible values.
        That is, incidence proportion and prevalence should be in the interval $[0, 1]$.
        While incidence proportion can theoretically be greater than 1, this would imply that some individuals were infected multiple times in a single time unit (day) which is not biologically plausible.
    \item Have sufficient parameters to incorporate all necessary covariates to correct for non-response bias (see \cref{E-biology-data:sec:positivity}).
    \item Computationally efficient.
        The model will be fit to granular data and therefore needs to be run within the SRS, with its computational limitations (see\todo{ref section on SRS computational limitations}).
        In particular, MCMC should be avoided because incorporating sufficient covariates to account for all groups led to a model that was too computationally expensive~\citePersonalComms{Koen Pouwels}.
\end{itemize}
